---
title: "GrupoBimboEDA"
author: "Anandan Balaji"
output: 
  pdf_document:
   toc: yes
   number_sections: yes
---

# Grupo Bimbo Exploratory Data Analysis

Objective: Predict the demand based on historical sales data.

## About Dataset & Data fields

The dataset is available at [www.kaagle.com](https://www.kaggle.com/c/grupo-bimbo-inventory-demand/data)

Some of the important data fields:

  *  Semana — Week number (From Thursday to Wednesday)
  *  Agencia_ID — Sales Depot ID
  *  Canal_ID — Sales Channel ID
  *  Ruta_SAK — Route ID (Several routes = Sales Depot)
  *  Cliente_ID — Client ID
  *  NombreCliente — Client name
  *  Producto_ID — Product ID
  *  NombreProducto — Product Name
  *  **Venta_uni_hoy — Sales unit this week** (integer)
  *  Venta_hoy — Sales this week (unit: pesos)
  *  Dev_uni_proxima — Returns unit next week (integer)
  *  Dev_proxima — Returns next week (unit: pesos)
  *  **Demanda_uni_equil — Adjusted Demand** (integer) (This is the target you will predict)

## Evaluation

  The evaulation metric for this competition is [Root Mean Squared Logarithmic Error](https://www.kaggle.com/wiki/RootMeanSquaredLogarithmicError).
  
## Train dataset

```{r}
library("data.table")
system.time(train <- fread("./new_train.csv", header = TRUE))

system.time(test <- fread("./test.csv", header = TRUE))

system.time(product <- fread("./producto_tabla.csv", header=TRUE))
```

```{r}
#structure of train
str(train)

## number of observations
nrow(train)

## get the weekly data - ie., number of transactions in that particular week.
table(train$Semana)

## get the demand info for every week
## tapply(X-vector, Index-variable, function)
##
tapply(train$Demanda_uni_equil, train$Semana, sum)

## number of unque products
length(unique(train$Producto_ID))

# which is the highest demand product
prod_results <- tapply(train$Demanda_uni_equil, train$Producto_ID, sum)
prod_results <- sort(prod_results, decreasing = TRUE)

highest_demand_prod <- prod_results[1]
highest_demand_prod

## the most popular product is

str(product)

product$NombreProducto[2425]
```


## Test Dataset

```{r}
# let's look at the test dataset

str(test)

# look at the week info
table(test$Semana)

#look at the products, are all of them available in training dataset?
train_prods <- unique(train$Producto_ID)

test_prods <- unique(test$Producto_ID)

# are the products equal
setequal(train_prods, test_prods)

# number of products equal
length(intersect(train_prods, test_prods))

# get the new products in test dataset
new_prods_in_test <- setdiff(test_prods, train_prods)

# number of new products in test dataset
length(new_prods_in_test)

## look at Agency ID (Sales Depot ID)
setequal(train$Agencia_ID, test$Agencia_ID)

## Channel ID (Sales chaneel ID)
setequal(train$Canal_ID, test$Canal_ID)

## Route ID ( Ruta_SAK)
setequal(train$Ruta_SAK, test$Ruta_SAK)

intersect_routes <- intersect(train$Ruta_SAK, test$Ruta_SAK)

#new routes in test
test_new_routes <- setdiff(test$Ruta_SAK, train$Ruta_SAK)

is.element(test_new_routes[1], test$Ruta_SAK)

## examine the client IDs
setequal(train$Cliente_ID, test$Cliente_ID)

new_test_clients <- setdiff(test$Cliente_ID, train$Cliente_ID)

# how many new clients?
length(new_test_clients)

is.element(new_test_clients[1], test$Cliente_ID)
```

## Plots

```{r}
library("ggplot2")
```

## Weekly Transactions.

```{r}
# per week, h
ggplot(train, aes(x = Semana)) + geom_histogram()
```

## Returns per week

```{r}
# returns per week

returns <- train[, .( total_returns = sum(Dev_uni_proxima)), by = Semana]

ggplot(returns, aes(x=Semana, y = total_returns)) + geom_line()

```

## Product Sales 

```{r}

# product wise sale (units) & return info.

prodReturn <- train[, .(sales = sum(Venta_uni_hoy), returns = sum(Dev_uni_proxima)), by = Producto_ID]

# product-wise sales (units)
ggplot(prodReturn, aes(x = Producto_ID, y = sales)) + geom_point()


```

## Product wise Sale vs Return 

```{r}
# product wise returns
ggplot(prodReturn, aes(x = Producto_ID, y = returns, color = returns)) + geom_point() +
  scale_color_gradient(low="blue", high="red")
```

## Demand Vs Sales

```{r}
# use just a fraction to plot as the table is huge!
# demand vs sales
# Demanda_uni_quil vs Venta_hoy

library("dplyr")
ggplot(train %>% sample_frac(0.05),aes(x = Venta_uni_hoy , y = Demanda_uni_equil)) +
  geom_point() +
  geom_smooth(method = "lm") +
  scale_x_continuous(name = "Sales") +
  scale_y_continuous(name = "Demand") +
  ggtitle("Demand Vs Sales (in terms of units)")
```


## Model
```{r}
#create the log. demand

train$log_demand <- log1p(train$Demanda_uni_equil)

# fit the training data
fit <- lm(Demanda_uni_equil ~ Venta_uni_hoy, data = train)
summary(fit)

fit2 <- lm(log_demand ~ Venta_uni_hoy, data = train)
summary(fit2)
```
